A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-07-20 demystify.tex
services

\paragraph{About Me} I am an associate professor in the CS dept at George Mason University (Mason). Before Mason, I was at the University of Nebraska-Lincoln (UNL). I have been in the PhD admission process at Mason and UNL for many years.  I also serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with the MS admission process---which is quite different than PhD). My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.

>>>>

\paragraph{About Me} I am an associate professor in the CS dept at George Mason University (Mason). Before Mason, I was at the University of Nebraska-Lincoln (UNL). I have been in the PhD admission process at Mason and UNL for many years. This allows me to have a good understanding of the process, the challenges students face, and what faculty are looking for. I also serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD). My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.

### 2024-07-20 faculty.tex
services

\chapter{Research}
\chapter{Obtaining Funding}
Getting funding is arguably the most important part of getting tenure.  Without funding, you can't pay your students, conference registration and travel, or even your own summer salary. However, writing grant proposals is very different from writing research papers. Thus, most new professors struggle with grant writing at first, but eventually get better at it.

@@

Many would (rightly) advise about being strategic about your services. I do not really generally this and just volunteer for services that I think it is my responsibility (e.g., being in PCs, though I never liked them much) and those that I enjoy (NSF panels, services to the depts). I enjoy services to my dept. and in many cases would take on tasks that no one wants to do because I feel that someone gotta do it and I can help out, e.g., various misc tasks in the dept.

@@

\section{Change in Departmental Leadership. Moving from UNL to GMU}

I joined UNL in 2016 and was very happy there until 2020, which is when the dept. changed its leadership and organization and lead to my decision to be on the job market again.

I do not want to get into details but below summarizes several things that happened

\begin{itemize}
\item 2019: four senior professors (in SE) left.  One of these 4 had became the chair of the dept and then left just after 2 months.  
\item 2020:  an even split among the faculty on moving to completely to the College of Engr.   Multiple senior people warn about this decision.  The new chair candidate only agrees to accept the job if the faculty moved to CoE, putting pressure on the move.
\item Salary inversion:  new chair offers new AP salary that is much higher than current AP, leading to jokes that current AP should just reapply to UNL.  This became a serious issue and many people at UNL use this as a reason on why they move.

>>>>

\chapter{Research}
Research here focuses on paper publication, which is a major component in getting your name out there and of course in your departmental evaluation.

\chapter{Obtaining Funding}
Getting funding is arguably the most important part of getting tenure.  Without funding, you can't pay your students, conference registration and travel, or even your own summer salary. However, writing grant proposals is very different from writing research papers. Thus, most new professors struggle with grant writing at first, but eventually get better at it.

@@

Many would (rightly) advise about being strategic about your services. I didn't really follow this and just volunteer for services that I think it is my responsibility (e.g., being in PCs, though I never liked them much) and those that I enjoy (NSF panels, services to the depts). I enjoy services to my dept. and in many cases would take on tasks that no one wants to do because I feel that someone gotta do it and I can help out, e.g., various misc tasks in the dept.

@@

\section{Change in Department. Moving from UNL to GMU}

I joined UNL in 2016 and was very happy there until 2020, when people start leaving and the dept. changed its leadership and organization.  All of these uncertainties and other issues lead to my decision to be on the job market again.

I will skip the details and just summarize several things that I didn't like.

\begin{itemize}
\item 2019: four senior professors (in SE) left.  One of these 4 was the dept chair (and he left just after 2 months of being chair). There were issues of upper admin didn't work well with the dept chair (and this issue has been going for years).
\item 2020: College of Engr pressure dept to move entirely to it. The new chair candidate only agrees to accept the job if the faculty moved to CoE, putting pressure on the move. Faculty are split on the decision and eventually moved with lots of mistrusts and concerns under CoE.
\item Salary inversion:  new chair offers new AP salary that is much higher than current AP, leading to jokes that current AP should just reapply to UNL.  This became a serious issue and many people at UNL use this as a reason on why they move. 
\item Senior faculty fighting on mailing list.
